### 2018-05-11 4_impl.tex
polish text a bit.

The main call for commencing the work on \porthos[2] was the need for processing real-world C programs, which, at first, requires the input language to be extended.
This implies the support not only for new syntactic structures of C language (such as the \texttt{switch} statement or the postfix increment operator \texttt{i++}), but also for its fundamental concepts and features (such as types, pointer arithmetic or first-order functions), which requires revision of the whole architecture of the tool.
Yet far from all of the C language has been supported (that, considering its complexity and numerous pitfalls, goes far beyond current thesis%
\footnote{For ensuring this, one merely has to look at existing C compilers, for instance, the open-source \tool{gcc} compiler, which uses the C parser written in more than 18.5 thousand lines of code (see \url{https://github.com/gcc-mirror/gcc/blob/master/gcc/c/c-parser.c})}%
), we consider the accomplished work as a step towards this.
%This applies not only to instances that simply constitute the syntactic sugar of C language (such as the \texttt{switch} statement or the prefix increment operator \texttt{++i}), but also to its fundamental concepts and features (such as pointer arithmetic or structures).
%However, the architecture of current version of \porthos{} does not allow to ... as it is ....
% Although and optimisation of the tool so that it is able to process real C programs in perspective.

@@

Therefore, while working on the new design of \porthos[2], we decided to clearly separate the high-level intermediate code representation (implemented as a recursive AST structure) from low-level event-based representation (implemented as an event-flow graph).
Such a modular architecture allows to support multiple input languages by parsing them and converting parsed syntax trees to a simplified AST. which, along with all other data-transfer objects (DTO), must be immutable, so that it is possible to guarantee the correctness of the program by controlling its invariants.
The immutability in \porthos[2] is implemented via \texttt{final} fields that are assigned by the immutable-object values (either a primitive type, or another immutable object, or an immutable collection provided by the library Guava by Google%
\footnote{Guava project repository: \url{https://github.com/google/guava/}}%
).

During the development of \porthos[2] we mainly followed the \textit{KISS~principle}, which can be exhaustively described in 17 Unix Rules of Eric Raymond~\cite{raymond2003art}.% which include the rule of modularity, the rule of clarity, transparency, extensibility, etc.
The following list summarises the main rules we followed during the development of \porthos[2]:

@@

The robustness of analysis is the main criterion of \porthos[2] as a verification of a bug-hunting tool.
This implies modelling the system with only over-approximating abstractions, which, however makes it more vulnerable to the combinatorial growth of the state space size, preserve the completeness of analysis.
As a robust and transparent tool, the \porthos[2] must adhere to the strategy of aborting its work on any unexpected state (for instance, if a parser failed to parse the string and recovery algorithm is not described).
%The modular architecture of \porthos[2] allow each its module to be tested independently.
%Usage of immutable data structures for all DTOs helps 
%immut: preservation of invariants
%The transparency, efficiency and extensibility are the necessary criteria of any good tool.

As its predecessor, \porthos[2] uses the open-source SMT-solver \tool{Z3} %
\footnote{The \tool{Z3} project repository: \url{https://github.com/Z3Prover/z3}}%
by Microsoft Research~\cite{de2008z3}. However, unlikely its predecessor, the \porthos[2] has an additional abstraction level \zformula{} (see Section~\ref{ch:impl:model:zformula}) that allow to use any other SMT-solver. %TODO: *should* allow?

@@

\begin{figure}%[H]
    \centering
  %draft=true
  \includegraphics[height=.99\textheight,keepaspectratio]{img/my/draw.io/general_arch.png}
  \caption{The general architecture of \porthos[2]}
  \label{fig:arch}
\end{figure}

The program takes as input the program to be analysed and one (the reachability analysis mode) or two (the portability analysis mode) memory models.
The parsed program syntax tree is then converted (Section~\ref{ch:impl:proc:y-constr}) to a program AST called \ytree{}%
\footnote{In order to avoid confusion between different internal representations, we prefix the names of elements of each internal representation with a letter. For instance, we picked the letter `Y' to denote the AST code representation as drawing of this letter resembles the tree branching; with letter `X' we prefix elements of the event-flow graph as the events are to be e\textbf{x}ecuted; and with letter `W' we prefix elements of the \textbf{w}eak memory model AST.}%
(Section~\ref{ch:impl:model:ytree}), which then is being preprocessed at the pre-compilation stage (Section~\ref{ch:impl:proc:x-pre-compiler}) in order to collect information necessary for the compilation.
The \ytree{} then is being compiled (Section~\ref{ch:impl:proc:x-compiler}) to an \xgraph{} representation (Section~\ref{ch:impl:model:xgraph}).
The compiled \xgraph{} then is being converted to an acyclic form (Section~\ref{ch:impl:proc:x-unroll}) in order to be encoded into a \zformula{} (Section~\ref{ch:impl:model:zformula}).
Apart from that, the memory-model constructor (Section~\ref{ch:impl:proc:w-constr}) constructs the abstract syntax tree of derived relations of the weak memory model \wmodel{} (Section~\ref{ch:impl:model:wmodel}).
%on the basis of the \xgraph{} in order to build the weak memory model \wmodel{} (Section~\ref{ch:impl:model:wmodel}).
Thereafter, \wmodel{} and acyclic \xgraph{} are encoded (Section~\ref{ch:impl:proc:z-encoder}) to a \zformula{} representation (a wrapper over an SMT-formula), which then is translated to an SMT-formula (Section~\ref{ch:impl:proc:smt-converter}), which then is solved by the SMT-solver (Section~\ref{ch:impl:proc:smt-solver}).

@@

Figure~\ref{fig:in_grammar_pts} represents the simplified grammar of of input language used by previous version of \porthos{} (the full \tool{ANTLR} grammar is available at Appendix~\ref{apx:grammar}.

@@

The input language parser used by \porthos{} suffered from several disadvantages.
Firstly, it contained the parser code inlined into directly the grammar, so that the grammar would serve as a template for the parser code (which is called semantic actions). Such a combining of two rich languages%
\footnote{by the term "reach" we mean "expressiveness": at least, the grammar of the language of semantic actions (i.e., java in our case) is Turing-complete.} %
makes the code hardly understandable, and, therefore, poorly maintainable. In \porthos[2], we clearly separated the parser (generated from the grammar file `\textit{<grammar>.g4}') from the converting the \tool{ANTLR} syntax tree to the AST, that is one for all  languages of an input program.

Secondly, in \porthos{}, the semantics of operations was defined syntactically (in \tool{ANTLR} grammar), whereas this should be performed by a separate module operating on the AST level%
\footnote{Most modern languages (including C++) allow the function overloading, thus requiring the type information to be available for the function resolution algorithm (see Section~\ref{ch:impl:proc:x-pre-compiler:type}).}%
, so that it does not require to change grammar for encoding the semantics of a new function. %adding support for a new function.

@@

The minor drawbacks of grammar used by \porthos{} include lack of operator associativity (expressions of the form \lstinline{1 + 2 * 3} could not be parsed), incorrectly (in terms of C) implemented grammar rule for the statement (the semicolon punctuator `\lstinline{;}' was implemented as the separator between two statements, whereas in C it serves as a statement terminator). Also, \porthos{} supports only the litmus-specific syntax for the variables initialisation, however, it allows only ini
tialisation of the shared variable and only by default value \lstinline{0}. The \porthos[2] supports arbitrary declaration to be performed in initialisation statement.

The \porthos[2] uses the C language grammar of proposed in the C11 standard~\cite{jtc2011sc22}, that was extended by litmus test-specific syntax such as initialisation and final-state assertion statements (the original \tool{ANTLR} grammar can be found in the official repository containing the collection of \tool{ANTLR v4} grammars%
\footnote{Repository path:~\url{https://github.com/antlr/grammars-v4}}).%
Current version of \porthos[2] does not recognise C processor directives (it ignores them), however, in future it can be extended to support them.

@@

The \ytree{} is compiled into the low-level event-based program representation called \textit{\xgraph{}}.
The mathematical structure of event-flow graph was discussed in Section~\ref{ch:wmm:event}.
%In general, \xgraph{} follows the structure described there, however it is constructed in several stages.
The nodes of the graph are events, and the edges are basic relations: the control-flow relation \po and the data-flow relations \co and \rf.
Hereinafter, we denote the \xgraph{} with only control-flow edges as \xgraph[CF], the \xgraph{} with only data-flow edges as \xgraph[CF].
The complete \xgraph{} is $\xgraph[CF+DF] = \xgraph[CF] \; \cup \; \xgraph[DF]$.
The UML diagraph on Figure~\ref{fig:class-diagrams:XEntity-interfaces} represents the hierarchy of main interfaces of the X-abstraction level.
%The \xgraph{} is constructed by the mutable graph builder. %TODO: name of builder -- after refactoring

@@

\begin{figure}[t]%[!b]%[H]
  \centering
  \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{img/my/class-diagrams/XMemoryUnit-m.png}
  \caption{The inheritance tree of \xgraph{} memory units}
  \label{fig:class-diagrams:XMemoryUnit}
\end{figure}

@@

Each memory unit has an \texttt{XType} associated with it.
The X-type is a symbolic representation of the C primitive type%
\footnote{Here we should note the \porthos[2] can eventually evolve to be able to analyse programs written in an OOP language (for instance, in C++). In this case, the \texttt{XType} will have more complex structure than a simple enumeration, which it has when we need to emulate only primitive types of C language. See more detailed discussion on input language type system in Section~\ref{ch:impl:proc:x-pre-compiler:type}.} %
that is easily convertible to an SMT-type (modelled as \texttt{ZType}).%TODO: add the class ZType

@@

\paragraph{Graph invariants}
\label{ch:impl:model:xgraph:invariants}
Once being constructed, the graph must conform the following requirements:

\begin{enumerate}
\item graph must have a single source with no ingoing edges, and two sinks of different kinds without outgoing edges,
\item graph must be connected,
\item each node can have either one or two direct control-flow successors,
\item only nodes of type \texttt{XComputationEvent} can have two direct control-flow successors,
%\item each node except the source node must have at least one parent
\item \textbf{todo: sth about data-flow edges}
\end{enumerate}

@@

@@

a simple set of recursively defined relations over ..

atoms: sets/basic relations

itemize elements

todo: rename everywhere assertions <- axioms

@@

in \porthos{} : ctx as an argument, calling ctx to create new clause. 
drawbacks:
- hard to debug
- only one solver
- non-safe way to construct formula (if ctx has changed, runtime excpetion)

in \porthos[2] we created the new abstraction layer \zformula[] that is translated to

recursive representation of SMT-formula

used as an abstraction btween

implemented as a simple wrapper over the z3 formula

typed: Expr, BoolExpr, ArithExpr, ...

@@

@@

the full \xgraph{} construction is performed in three stages.
First, the \ytree{} is compiled to a \textit{cyclic} control-flow event-based graph \xgraph[CF].
Then, this graph is unrolled to an \textit{acyclic} control-flow event-based graph \xgraphU[CF].
After that, the compiler is able to perform the data-flow analysis and produce the full event-based graph \xgraphU[CF+DF], which remains to be \textit{CF-acyclic} (no cycles among control-flow edges).

@@

@@

The input memory model language \cat{} is discussed in Section~\ref{ch:wmm:cat}.
The \tool{ANTLR} grammar for the \cat{} language was extracted from the parser used by \tool{herd} tool%
\footnote{\tool{herd} project repository: \url{https://github.com/herd/herdtools7}} %
written in OCaml.
%TODO: upload the grammar to github
%The input memory model 
\textbf{TODO: add reference to the grammar file in github repository OR include a short version as an appendix}

@@

@@

@@

The \xgraph{} abstract machine model disposes infinite number of memory units, both local and shared (see Section~\ref{ch:impl:model:xgraph:mem}).
The X-compiler accesses all memory units via the \texttt{XMemoryManager}, which by the end of pre-compilation stage is already initialised (has registered all shared memory units).
However, the memory manager is a stateful component of the compiler as it offers the methods for declaring and removing local memory units dynamically at the compilation time.
At each time of the compilation process, the \texttt{XMemoryManager} can resolve the memory unit by its name.
Following the C standard, local memory units have higher priority over global ones.
Since the C language allows the variables that have the same name to be declared in nested contexts, the \texttt{XMemoryManager} carries the block-context stack. %TODO: refactor XMemoryManager!

>>>>

The main call for commencing the work on \porthos[2] was the need for processing real-world C programs, which, at first, requires the input language to be extended.
This implies the support not only for new syntactic structures of C language (such as the \texttt{switch} statement or the postfix increment operator \texttt{i++}), but also for its fundamental concepts and features (such as types, pointer arithmetic or first-order functions), which requires revision of the whole architecture of the tool.
Yet far from all of the C language has been supported (that, considering its complexity and numerous pitfalls, goes far beyond current thesis%
%
\footnote{For ensuring this, one merely has to look at existing C compilers, for instance, the open-source \tool{gcc} compiler, which uses the C parser written in more than 18.5 thousand lines of code (see \url{https://github.com/gcc-mirror/gcc/blob/master/gcc/c/c-parser.c})}%
%
), we consider the accomplished work as a step towards it.
%This applies not only to instances that simply constitute the syntactic sugar of C language (such as the \texttt{switch} statement or the prefix increment operator \texttt{++i}), but also to its fundamental concepts and features (such as pointer arithmetic or structures).
%However, the architecture of current version of \porthos{} does not allow to ... as it is ....
% Although and optimisation of the tool so that it is able to process real C programs in perspective.

@@

Therefore, while working on the new design of \porthos[2], we decided to clearly separate the high-level intermediate code representation (implemented as a recursive AST structure) from low-level event-based representation (implemented as an event-flow graph).
Such a modular architecture allows to support multiple input languages by parsing them and converting parsed syntax trees to a simplified AST. which, along with all other data-transfer objects (DTO), must be immutable, so that it is possible to guarantee the correctness of the program by controlling its invariants.
The immutability in \porthos[2] is implemented via \texttt{final} fields that are assigned by the immutable-object values (either a primitive type, or another immutable object, or an immutable collection provided by the library Guava by Google%
%
\footnote{Guava project repository: \url{https://github.com/google/guava/}}%
%
).

During the development of \porthos[2] we mainly followed the \textit{KISS~principle}, which can be exhaustively described in 17 Unix Rules of Eric Raymond~\cite{raymond2003art}.
The following list summarises the main rules we followed during the development of \porthos[2]:

@@

Robustness of the analysis is the main criterion of \porthos[2] as a verification tool.
This implies modelling the system with usage of only over-approximating abstractions 
However it makes the analysis more sensitive to the combinatorial explosion, it preserves the completeness of analysis, which is necessary for a model-checking tool.
As a robust and transparent tool, \porthos[2] must adhere to the strategy of aborting its work on any unexpected state (for instance, if a parser failed to parse the string and recovery algorithm is not described).
%The modular architecture of \porthos[2] allow each its module to be tested independently.
%Usage of immutable data structures for all DTOs helps 
%immut: preservation of invariants
%The transparency, efficiency and extensibility are the necessary criteria of any good tool.

As its predecessor, \porthos[2] uses the open-source SMT-solver \tool{Z3} %
%
\footnote{The \tool{Z3} project repository: \url{https://github.com/Z3Prover/z3}}%
%
by Microsoft Research~\cite{de2008z3}. However, unlikely its predecessor, the \porthos[2] has an additional abstraction level \zformula{} (see Section~\ref{ch:impl:model:zformula}) that allow to use any other SMT-solver. %TODO: *should* allow?

@@

\begin{figure}
  \centering
  %draft=true
  \includegraphics[height=.99\textheight,keepaspectratio]{img/my/draw.io/general_arch.png}
  \caption{The general architecture of \porthos[2]}
  \label{fig:arch}
\end{figure}

The program takes as input the program to be analysed and one (the reachability analysis mode) or two (the portability analysis mode) memory models.
The parsed program syntax tree is then converted (Section~\ref{ch:impl:proc:y-constr}) to a program AST called \ytree{}%
%
\footnote{In order to avoid confusion between different internal representations, we prefix the names of elements of each internal representation with a letter. For instance, we picked the letter `Y' to denote the AST code representation as drawing of this letter resembles the tree branching; with letter `X' we prefix elements of the event-flow graph as the events are to be e\textbf{x}ecuted; and with letter `W' we prefix elements of the \textbf{w}eak memory model AST.}%
%
(Section~\ref{ch:impl:model:ytree}), which then is being preprocessed at the pre-compilation stage (Section~\ref{ch:impl:proc:x-pre-compiler}) in order to collect information necessary for the compilation.
The \ytree{} then is being compiled (Section~\ref{ch:impl:proc:x-compiler}) to an \xgraph{} representation (Section~\ref{ch:impl:model:xgraph}).
The compiled \xgraph{} then is being converted to an acyclic form (Section~\ref{ch:impl:proc:x-unroll}) in order to be encoded into a \zformula{} (Section~\ref{ch:impl:model:zformula}).
Apart from that, the memory-model constructor (Section~\ref{ch:impl:proc:w-constr}) constructs the abstract syntax tree of derived relations of the weak memory model \wmodel{} (Section~\ref{ch:impl:model:wmodel}).
%on the basis of the \xgraph{} in order to build the weak memory model \wmodel{} (Section~\ref{ch:impl:model:wmodel}).
Thereafter, \wmodel{} and acyclic \xgraph{} are encoded (Section~\ref{ch:impl:proc:z-encoder}) to a \zformula{} representation (a wrapper over an SMT-formula), which then is translated to an SMT-formula (Section~\ref{ch:impl:proc:smt-converter}), which then is solved by the SMT-solver (Section~\ref{ch:impl:proc:smt-solver}).

@@

@@

Figure~\ref{fig:in_grammar_pts} represents the simplified grammar of of input language used by previous version of \porthos{} (the full \tool{ANTLR} grammar is available at Appendix~\ref{apx:grammar}.
The input language parser used by \porthos{} suffered from several disadvantages.
Firstly, it contained the parser code inlined into directly the grammar, so that the grammar would serve as a template for the parser code (which is called semantic actions). Such a combining of two expressive languages makes the code hardly understandable and, therefore, poorly maintainable.
In \porthos[2], we clearly separated the parser (generated from the grammar file `\textit{<grammar>.g4}') from the converting the \tool{ANTLR} syntax tree to the AST, that is one for all languages of an input program.

Secondly, in \porthos{}, the semantics of operations was defined syntactically (in \tool{ANTLR} grammar), whereas this should be performed by a separate module operating on the AST level%
%
\footnote{Most modern languages (including C++) allow the function overloading, thus requiring the type information to be available for the function resolution algorithm (see Section~\ref{ch:impl:proc:x-pre-compiler:type}).}%
%
, so that it does not require to change grammar for encoding the semantics of a new function. %adding support for a new function.

@@

The minor drawbacks of grammar used by \porthos{} include lack of operator associativity (expressions of the form `\lstinline{1 + 2 * 3}' could not be parsed), incorrectly (in terms of C) implemented grammar rule for the statement (the semicolon punctuator `\lstinline{;}' was implemented as the separator between two statements, whereas in C it serves as a statement terminator). Also, \porthos{} supports only the litmus-specific syntax for the variables initialisation, however, it allows only ini
tialisation of the shared variable and only by default value \lstinline{0}. \porthos[2] supports arbitrary declaration to be performed in initialisation statement.

\porthos[2] uses the C language grammar of proposed in the C11 standard~\cite{jtc2011sc22}, that was extended by litmus test-specific syntax such as initialisation and final-state assertion statements (the original \tool{ANTLR} grammar can be found in the official repository containing the collection of \tool{ANTLR v4} grammars%
%
\footnote{Repository path:~\url{https://github.com/antlr/grammars-v4}}). %
%
Current version of \porthos[2] does not recognise C processor directives (it ignores them), however, in future it can be extended to support them.

@@

The \ytree{} is compiled into the low-level event-based program representation called \textit{\xgraph{}}.
The mathematical structure of event-flow graph was discussed in Section~\ref{ch:wmm:event}.
%In general, \xgraph{} follows the structure described there, however it is constructed in several stages.
The nodes of the graph are events, and the edges are basic relations: the control-flow relation \po and the data-flow relations \co and \rf.
Hereinafter, we denote the \xgraph{} with only control-flow edges as \xgraph[CF], the \xgraph{} with only data-flow edges as \xgraph[CF].
The complete \xgraph{} is $\xgraph[CF+DF] = \xgraph[CF] \; \cup \; \xgraph[DF]$.
The UML diagram on Figure~\ref{fig:class-diagrams:XEntity-interfaces} represents the hierarchy of main interfaces of the X-abstraction level.
%The \xgraph{} is constructed by the mutable graph builder. %TODO: name of builder -- after refactoring

@@

%TODO!!! Remove watermark from the pictures!
\begin{figure}[t]%[!b]%[H]
  \centering
  \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{img/my/class-diagrams/XMemoryUnit-m.png}
  \caption{The inheritance tree of \xgraph{} memory units}
  \label{fig:class-diagrams:XMemoryUnit}
\end{figure}

@@

Each memory unit has an \texttt{XType} associated with it.
The X-type is a symbolic representation of the C primitive type%
%
\footnote{Here we should note that \porthos[2] can eventually evolve to be able to analyse programs written in an OOP language (for instance, in C++). In this case, the \texttt{XType} will have more complex structure than a simple enumeration, which it has when we need to emulate only primitive types of C language. See more detailed discussion on input language type system in Section~\ref{ch:impl:proc:x-pre-compiler:type}.} %TODO!!!!! Check that XType is really a simple enumeration
%
that is easily convertible to an SMT-type (modelled as \texttt{ZType}).%TODO: add the class ZType

@@

\paragraph{Graph invariants}
\label{ch:impl:model:xgraph:invariants}
Once being constructed, the graph must conform the following requirements (see also Section~\ref{ch:port:enc:cf}):

\begin{enumerate}[noitemsep]
\item the graph must have a single source with no ingoing edges, and two sinks of different kinds without outgoing edges,
\item the graph must be connected,
\item each node of the graph can have either one or two direct control-flow successors,
\item only nodes of type \texttt{XComputationEvent} can have two direct control-flow successors,
%\item each node except the source node must have at least one parent
\item a \co-edge connects two writes, an \rf-relation edge connects a write and a read,
\item all write-event nodes except initial write-nodes must have exactly one \co-predecessor
\item all write-event nodes except final write-nodes must have exactly one \co-successor.
%TODO: \item \textbf{TODO: more?}
\end{enumerate}

@@

<TODO: finish it later after fixing the code>

@@

A \wmodel{} represents a recursively defined AST of relations defined by the memory model and an assertion expression.
The atomic elements of \wmodel{} are the basic relations (\po, \rf and \co; see Section~\ref{ch:wmm:model:relations}) and sets of events ($\mathbb{R}$, $\mathbb{W}$, $\mathbb{IW}$, etc.; see Section~\ref{ch:wmm:model:events}).
The expressions of \wmodel{} are unary (such as complement, transitive closure, etc.) and binary operations (such as union, intersection, etc.) over relations or sets of events; see Section~\ref{ch:wmm:cat}.

<todo: enumerate elements>

<carries the origin -- from cat file>

%todo: rename everywhere assertions <- axioms

@@

The \zformula{} representation is a wrapper for an SMT-formula used as an additional abstraction level used to increase the transparency of the architecture, simplify the debugging process and ease the support of different SMT-solvers.
As all other internal representations, the \zformula{} is implemented as an immutable data structure.

Following the structure of a Z3-formula, the atomic elements of a \zformula{} are variables, and the expressions are

reference to Z3-tutorial AND

All expressions and variables of a \zformula{} are typed.
Basically, we distinguish \textit{boolean} and \textit{numeric} (integer, floating point, <else?>) data type

%typed: Expr, BoolExpr, ArithExpr, ...

@@

<todo: enumerate elements>

@@

The construction of the \xgraph{} is performed in three stages.
First, the \ytree{} is compiled to a \textit{cyclic} control-flow event-based graph \xgraph[CF].
Then, this graph is unrolled to an \textit{acyclic} control-flow event-based graph \xgraphU[CF].
After that, the compiler is able to perform the data-flow analysis and produce the full event-based graph \xgraphU[CF+DF], which remains to be \textit{CF-acyclic} (no cycles among control-flow edges).

@@

Construction of immutable data structures is performed by units that follow the \textit{builder} pattern.
<TODO: SAY MORE: validation of ht einput elements during construction, maybe verif after building>

@@

The input memory model language \cat{} is discussed in Section~\ref{ch:wmm:cat}.
The \tool{ANTLR} grammar for the \cat{} language was extracted from the parser used by \tool{herd} tool%
%
\footnote{\tool{herd} project repository: \url{https://github.com/herd/herdtools7}} %
%
written in OCaml.
%TODO: upload the grammar to github
%The input memory model 
\textbf{TODO: add reference to the grammar file in github repository OR include a short version as an appendix}

@@

builder: filling in + verification

@@

The construction of a \zformula{} is performed by the \texttt{ZformulaBuilder}.%TODO: check this class existence

@@

The \xgraph{} abstract machine model disposes infinitely many memory units, both local and shared (see Section~\ref{ch:impl:model:xgraph:mem}).
The X-compiler accesses all memory units via the \texttt{XMemoryManager}, which by the end of pre-compilation stage is already initialised (has registered all shared memory units).
However, the memory manager is a stateful component of the compiler as it offers the methods for declaring and removing local memory units dynamically at the compilation time.
At each time of the compilation process, the \texttt{XMemoryManager} can resolve the memory unit by its name.
Following the C standard, local memory units have higher priority over global ones.
Since the C language allows the variables that have the same name to be declared in nested contexts, the \texttt{XMemoryManager} carries the block-context stack. %TODO: refactor XMemoryManager!
